---
title: "AFLxMargin"
author: "Josh Ryan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(fitzRoy)
library(dplyr)
library(purrr)
```

# Introduction

AFLxMargin is a multiple regression model used to predict the margin of AFL
games. Past AFL data is scraped from the web using the fitzRoy package, and
used as the predictors of the model. The predictors of the model are:

* The margin of the last match up between the two teams
* The number of wins each team has in their last 5 games
* The ladder positions of the teams

The model always predicts the margin from the perspective of the home team in
the match up. As a result, it can be thought that the home ground advantage is
also considered as an implicit predictor through the intercept of the model.

# The function for the model
Given below is the R code for the function of the model, which will produce a
prediction for a match given a round number and a match id, corresponding
to where that match lands in the fixture (e.g. 2nd match will have id = 2).
Note that the function as given below is currently card coded to only account for
the 2023 season.

```{r}
# Function for parsing ladder form string
parse_form_string <- function(string) {
  split_string <- strsplit(string, "")[[1]]
  
  result <- 0
  
  for (letter in split_string) {
    if (letter == "W") {
      result <- result + 1
    } else if (letter == "D") {
      result <- result + 0.5
    }
  }
  
  return( result)
}

# Function for getting last n characters of a string
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

# Main model function
afl_x_margin <- function(round_number, match_id) {
  round_info <- fetch_fixture(2023, round = round_number) %>% 
    select(home.team.abbreviation, away.team.abbreviation) %>% 
    mutate(id = row_number()) %>% 
    rename("home_team" = "home.team.abbreviation",
           "away_team" = "away.team.abbreviation")
  
  chosen_match_info <- round_info %>% filter(id == match_id) %>% select(home_team, away_team)
  
  # Fetch previous x matches
  previous_round_info <- data.frame()
  
  for (round in 6:round_number-1) { # starts at 6 to allow for past 5 game form
    round_data <- fetch_fixture(2023, round = round) %>% 
      select(home.team.abbreviation, away.team.abbreviation,
             home.score.totalScore, away.score.totalScore) %>% 
      rename("home_team" = "home.team.abbreviation",
             "away_team" = "away.team.abbreviation",
             "home_score" = "home.score.totalScore",
             "away_score" = "away.score.totalScore") %>% 
      mutate(round_no = round)
    
    previous_round_info <- bind_rows(previous_round_info, round_data)
  }
  
  # Get margin from scores
  previous_round_info <- previous_round_info %>%
    mutate(margin = home_score - away_score)
  
  # Fetch ladder positions
  ladder_positions <- fetch_ladder() %>%
    select(team.abbreviation, position)
  
  # Bind ladder positions with previous_round_info
  previous_round_info <- previous_round_info %>%
    left_join(ladder_positions, by = c("home_team" = "team.abbreviation")) %>%
    rename(home_position = position) %>%
    left_join(ladder_positions, by = c("away_team" = "team.abbreviation")) %>%
    rename(away_position = position)
  
  # Fetch scores of last match-up between teams
  season_results <- fetch_results(2023)
  last_season_results <- fetch_results(2022)
  for (row in 1:nrow(previous_round_info)) {
    season_matchup_info <- season_results %>%  arrange(desc(row_number())) %>%
      filter(match.homeTeam.abbr == previous_round_info[row, "home_team"] |
               match.homeTeam.abbr == previous_round_info[row, "away_team"],
             match.awayTeam.abbr == previous_round_info[row, "away_team"] |
               match.awayTeam.abbr == previous_round_info[row, "home_team"]) %>% 
      select(homeTeamScore.matchScore.totalScore,
             awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)
    
    last_season_matchup_info <- last_season_results %>% arrange(desc(row_number())) %>%
      filter(match.homeTeam.abbr == previous_round_info[row, "home_team"] |
               match.homeTeam.abbr == previous_round_info[row, "away_team"],
             match.awayTeam.abbr == previous_round_info[row, "away_team"] |
               match.awayTeam.abbr == previous_round_info[row, "home_team"]) %>% 
      select(homeTeamScore.matchScore.totalScore,
             awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)
    
    matchup_info <- bind_rows(season_matchup_info, last_season_matchup_info)
    
    if (previous_round_info[row, "home_team"] == matchup_info[2, "match.homeTeam.abbr"]) {
      previous_round_info[row, "previous_match_home_score"] <- matchup_info[2, "homeTeamScore.matchScore.totalScore"]
      previous_round_info[row, "previous_match_away_score"] <- matchup_info[2, "awayTeamScore.matchScore.totalScore"]
    } else {
      previous_round_info[row, "previous_match_home_score"] <- matchup_info[2, "awayTeamScore.matchScore.totalScore"]
      previous_round_info[row, "previous_match_away_score"] <- matchup_info[2, "homeTeamScore.matchScore.totalScore"]
    }
  }
  
  # Get margin from total scores
  previous_round_info <- previous_round_info %>%
    mutate(last_match_margin = previous_match_home_score - previous_match_away_score)
  
  # Select only necessary columns
  previous_round_info <- previous_round_info %>% 
    select(home_team, away_team, margin, home_position, away_position, last_match_margin, round_no)
  
  # Get wins in last 5 games for team being predicted
  for (row in 1:nrow(previous_round_info)) {
    previous_round_info[row, "predicted_form"] <- fetch_ladder(2023, round_number = previous_round_info[row, "round_no"]) %>%
      filter(team.abbreviation == previous_round_info[row, "home_team"]) %>% 
      select(form)
  }
  
  for (row in 1:nrow(previous_round_info)) {
    previous_round_info[row, "predicted_form"] = previous_round_info[row, "predicted_form"] %>% 
      substrRight(5)
  }
  
  previous_round_info <- previous_round_info %>%
    mutate(wins_in_last_5 = map_chr(predicted_form, parse_form_string))
  
  previous_round_info <- previous_round_info %>%
    select(-predicted_form)
  
  # Get wins in last 5 games for opponent team
  for (row in 1:nrow(previous_round_info)) {
    previous_round_info[row, "opponent_form"] <- fetch_ladder(2023, round_number = previous_round_info[row, "round_no"]) %>%
      filter(team.abbreviation == previous_round_info[row, "away_team"]) %>% 
      select(form)
  }
  
  for (row in 1:nrow(previous_round_info)) {
    previous_round_info[row, "opponent_form"] = previous_round_info[row, "opponent_form"] %>% 
      substrRight(5)
  }
  
  previous_round_info <- previous_round_info %>%
    mutate(opponent_wins_in_last_5 = map_chr(opponent_form, parse_form_string))
  
  previous_round_info <- previous_round_info %>%
    select(-opponent_form)
  
  previous_round_info$wins_in_last_5 <- 
    as.numeric(previous_round_info$wins_in_last_5)
  previous_round_info$opponent_wins_in_last_5 <- 
    as.numeric(previous_round_info$opponent_wins_in_last_5)
  
  # Get all player stats for the season up to round 16
  season_stats <- data.frame()
  for (round in 1:round_number-1) {
    round_stats <- fetch_player_stats(season = 2023, round_number = round)
    season_stats <- bind_rows(season_stats, round_stats)
  }
  
  # Get sum of ranking points of lineup from each home player's last 5 games
  for (row in 1:nrow(previous_round_info)) {
    home_lineup <- fetch_lineup(2023, previous_round_info[row, "round_no"]) %>%
      filter(teamAbbr == previous_round_info[row, "home_team"])
    
    home_rating <- 0
    
    for (prow in 1:nrow(home_lineup)) { # iterate over each player
      if (home_lineup[prow, "position"] %>% pull() != "EMERG") {
        player <- home_lineup[prow, ]
        
        pstats <- season_stats %>% 
          filter(player.player.player.playerId ==
                   player[1, "player.playerId"] %>% pull()) %>% 
          tail(5)
        
        home_rating <- home_rating + sum(pstats$dreamTeamPoints)
      }
    }
    
    previous_round_info[row, "home_rating"] <- home_rating
  }
  
  # Get sum of ranking points of lineup from each away player's last 5 games
  for (row in 1:nrow(previous_round_info)) {
    away_lineup <- fetch_lineup(2023, previous_round_info[row, "round_no"]) %>%
      filter(teamAbbr == previous_round_info[row, "away_team"])
    
    away_rating <- 0
    
    for (prow in 1:nrow(away_lineup)) { # iterate over each player
      if (away_lineup[prow, "position"] %>% pull() != "EMERG") {
        player <- away_lineup[prow, ]
        
        pstats <- season_stats %>% 
          filter(player.player.player.playerId ==
                   player[1, "player.playerId"] %>% pull()) %>% 
          tail(5)
        
        away_rating <- away_rating + sum(pstats$dreamTeamPoints)
      }
    }
    
    previous_round_info[row, "away_rating"] <- away_rating
  }
  
  model <- lm(margin ~ home_position + away_position + last_match_margin + 
                wins_in_last_5 + opponent_wins_in_last_5 + home_rating + 
                away_rating, data = previous_round_info)
  
  # Fetch ladder positions
  ladder_positions <- fetch_ladder() %>%
    select(team.abbreviation, position)
  
  # Bind ladder positions with chosen_match_info
  chosen_match_info <- chosen_match_info %>%
    left_join(ladder_positions, by = c("home_team" = "team.abbreviation")) %>%
    rename(home_position = position) %>%
    left_join(ladder_positions, by = c("away_team" = "team.abbreviation")) %>%
    rename(away_position = position)
  
  # Fetch scores of last match-up between teams
  season_matchup_info <- fetch_results() %>%  arrange(desc(row_number())) %>%
    filter(match.homeTeam.abbr == chosen_match_info[1, "home_team"] %>% pull |
             match.homeTeam.abbr == chosen_match_info[1, "away_team"] %>% pull,
           match.awayTeam.abbr == chosen_match_info[1, "away_team"] %>% pull |
             match.awayTeam.abbr == chosen_match_info[1, "home_team"] %>% pull) %>% 
    select(homeTeamScore.matchScore.totalScore,
           awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)
  
  last_season_matchup_info <- fetch_results(2022) %>% arrange(desc(row_number())) %>%
    filter(match.homeTeam.abbr == chosen_match_info[1, "home_team"] %>% pull |
             match.homeTeam.abbr == chosen_match_info[1, "away_team"] %>% pull,
           match.awayTeam.abbr == chosen_match_info[1, "away_team"] %>% pull |
             match.awayTeam.abbr == chosen_match_info[1, "home_team"] %>% pull) %>% 
    select(homeTeamScore.matchScore.totalScore,
           awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)
  
  matchup_info <- bind_rows(season_matchup_info, last_season_matchup_info)
  
  if (chosen_match_info[1, "home_team"] == matchup_info[1, "match.homeTeam.abbr"]) {
    chosen_match_info[1, "previous_match_home_score"] <- matchup_info[1, "homeTeamScore.matchScore.totalScore"]
    chosen_match_info[1, "previous_match_away_score"] <- matchup_info[1, "awayTeamScore.matchScore.totalScore"]
  } else {
    chosen_match_info[1, "previous_match_home_score"] <- matchup_info[1, "awayTeamScore.matchScore.totalScore"]
    chosen_match_info[1, "previous_match_away_score"] <- matchup_info[1, "homeTeamScore.matchScore.totalScore"]
  }
  
  # Get margin from total scores
  chosen_match_info <- chosen_match_info %>%
    mutate(last_match_margin = previous_match_home_score - previous_match_away_score)
  
  # Get wins in last 5 games for home team
  chosen_match_info[1, "home_form"] <- fetch_ladder(2023) %>%
    filter(team.abbreviation == chosen_match_info[1, "home_team"] %>% pull()) %>% 
    select(form)
  
  chosen_match_info[1, "home_form"] = chosen_match_info[1, "home_form"] %>% 
    substrRight(5)
  
  chosen_match_info <- chosen_match_info %>%
    mutate(wins_in_last_5 = map_chr(home_form, parse_form_string))
  
  # Get wins in last 5 games for away team
  chosen_match_info[1, "away_form"] <- fetch_ladder(2023) %>%
    filter(team.abbreviation == chosen_match_info[1, "away_team"] %>% pull()) %>% 
    select(form)
  
  chosen_match_info[1, "away_form"] = chosen_match_info[1, "away_form"] %>% 
    substrRight(5)
  
  chosen_match_info <- chosen_match_info %>%
    mutate(opponent_wins_in_last_5 = map_chr(away_form, parse_form_string))
  
  chosen_match_info <- chosen_match_info %>%
    select(-previous_match_home_score, -previous_match_away_score, 
           -home_form, -away_form)
  
  chosen_match_info$wins_in_last_5 <- 
    as.numeric(chosen_match_info$wins_in_last_5)
  chosen_match_info$opponent_wins_in_last_5 <- 
    as.numeric(chosen_match_info$opponent_wins_in_last_5)
  
  # Get all player stats for the season up to round 16
  season_stats <- data.frame()
  for (round in 1:round_number-1) {
    round_stats <- fetch_player_stats(season = 2023, round_number = round)
    season_stats <- bind_rows(season_stats, round_stats)
  }
  
  # Get sum of ranking points of lineup from each home player's last 5 games
  home_lineup <- fetch_lineup(2023, 16) %>%
    filter(teamAbbr == chosen_match_info[1, "home_team"] %>% pull())
  
  home_rating <- 0
  
  for (prow in 1:nrow(home_lineup)) { # iterate over each player
    if (home_lineup[prow, "position"] %>% pull() != "EMERG") {
      player <- home_lineup[prow, ]
      
      pstats <- season_stats %>% 
        filter(player.player.player.playerId ==
                 player[1, "player.playerId"] %>% pull()) %>% 
        tail(5)
      
      home_rating <- home_rating + sum(pstats$dreamTeamPoints)
    }
  }
  
  chosen_match_info[1, "home_rating"] <- home_rating
  
  # Get sum of ranking points of lineup from each away player's last 5 games
  
  away_lineup <- fetch_lineup(2023, 16) %>%
    filter(teamAbbr == chosen_match_info[1, "away_team"] %>% pull())
  
  away_rating <- 0
  
  for (prow in 1:nrow(away_lineup)) { # iterate over each player
    if (away_lineup[prow, "position"] %>% pull() != "EMERG") {
      player <- away_lineup[prow, ]
      
      pstats <- season_stats %>% 
        filter(player.player.player.playerId ==
                 player[1, "player.playerId"] %>% pull()) %>% 
        tail(5)
      
      away_rating <- away_rating + sum(pstats$dreamTeamPoints)
    }
  }
  
  chosen_match_info[1, "away_rating"] <- away_rating
  
  model_result <- predict(model, newdata = chosen_match_info)
  
  return(model_result)
}
```

Here is an example output of the model for a match between Essendon and Adelaide
in Round 17 of the 2023 season, where Essendon are the home team:

```{r}
afl_x_margin(17, 8)
```

This result should be interpreted as an expected margin of 9.83 in Essendon's
favor.

# Hard-coded example
The following is a hard coded example of the model, used to predict the result
of a match between Geelong and North Melbourne in round 17 of the 2023 season,
where Geelong is the home team.

Function for parsing ladder form string:
```{r}
parse_form_string <- function(string) {
  split_string <- strsplit(string, "")[[1]]
  
  result <- 0
  
  for (letter in split_string) {
    if (letter == "W") {
      result <- result + 1
    } else if (letter == "D") {
      result <- result + 0.5
    }
  }
  
  return( result)
}
```

Function for getting last n characters of a string:
```{r}
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
```

Scraping fixture information for given round:
```{r}
round_info <- fetch_fixture(2023, round = 17) %>% 
  select(home.team.abbreviation, away.team.abbreviation) %>% 
  mutate(id = row_number()) %>% 
  rename("home_team" = "home.team.abbreviation",
         "away_team" = "away.team.abbreviation")
```

Get teams for individual match:
```{r}
chosen_match_info <- round_info %>% filter(id == 7) %>% select(home_team, away_team)
```

Fetch predictor data for previous 5 matches from perspective of home team:
```{r}
# Fetch previous x matches
previous_round_info <- data.frame()

for (round in 6:16) {
  round_data <- fetch_fixture(2023, round = round) %>% 
    select(home.team.abbreviation, away.team.abbreviation,
           home.score.totalScore, away.score.totalScore) %>% 
    rename("home_team" = "home.team.abbreviation",
           "away_team" = "away.team.abbreviation",
           "home_score" = "home.score.totalScore",
           "away_score" = "away.score.totalScore") %>% 
    mutate(round_no = round)
  
  previous_round_info <- bind_rows(previous_round_info, round_data)
}

# Get margin from scores
previous_round_info <- previous_round_info %>%
  mutate(margin = home_score - away_score)

# Fetch ladder positions
ladder_positions <- fetch_ladder() %>%
  select(team.abbreviation, position)

# Bind ladder positions with previous_round_info
previous_round_info <- previous_round_info %>%
  left_join(ladder_positions, by = c("home_team" = "team.abbreviation")) %>%
  rename(home_position = position) %>%
  left_join(ladder_positions, by = c("away_team" = "team.abbreviation")) %>%
  rename(away_position = position)

# Fetch scores of last match-up between teams
season_results <- fetch_results(2023)
last_season_results <- fetch_results(2022)
for (row in 1:nrow(previous_round_info)) {
  season_matchup_info <- season_results %>%  arrange(desc(row_number())) %>%
    filter(match.homeTeam.abbr == previous_round_info[row, "home_team"] |
           match.homeTeam.abbr == previous_round_info[row, "away_team"],
           match.awayTeam.abbr == previous_round_info[row, "away_team"] |
           match.awayTeam.abbr == previous_round_info[row, "home_team"]) %>% 
    select(homeTeamScore.matchScore.totalScore,
           awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)
  
  last_season_matchup_info <- last_season_results %>% arrange(desc(row_number())) %>%
    filter(match.homeTeam.abbr == previous_round_info[row, "home_team"] |
           match.homeTeam.abbr == previous_round_info[row, "away_team"],
           match.awayTeam.abbr == previous_round_info[row, "away_team"] |
           match.awayTeam.abbr == previous_round_info[row, "home_team"]) %>% 
    select(homeTeamScore.matchScore.totalScore,
           awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)
  
  matchup_info <- bind_rows(season_matchup_info, last_season_matchup_info)
  
  if (previous_round_info[row, "home_team"] == matchup_info[2, "match.homeTeam.abbr"]) {
    previous_round_info[row, "previous_match_home_score"] <- matchup_info[2, "homeTeamScore.matchScore.totalScore"]
    previous_round_info[row, "previous_match_away_score"] <- matchup_info[2, "awayTeamScore.matchScore.totalScore"]
  } else {
    previous_round_info[row, "previous_match_home_score"] <- matchup_info[2, "awayTeamScore.matchScore.totalScore"]
    previous_round_info[row, "previous_match_away_score"] <- matchup_info[2, "homeTeamScore.matchScore.totalScore"]
  }
}

# Get margin from total scores
previous_round_info <- previous_round_info %>%
  mutate(last_match_margin = previous_match_home_score - previous_match_away_score)

# Select only necessary columns
previous_round_info <- previous_round_info %>% 
  select(home_team, away_team, margin, home_position, away_position, last_match_margin, round_no)

# Get wins in last 5 games for team being predicted
for (row in 1:nrow(previous_round_info)) {
  previous_round_info[row, "predicted_form"] <- fetch_ladder(2023, round_number = previous_round_info[row, "round_no"]) %>%
    filter(team.abbreviation == previous_round_info[row, "home_team"]) %>% 
    select(form)
}

for (row in 1:nrow(previous_round_info)) {
  previous_round_info[row, "predicted_form"] = previous_round_info[row, "predicted_form"] %>% 
    substrRight(5)
}

previous_round_info <- previous_round_info %>%
  mutate(wins_in_last_5 = map_chr(predicted_form, parse_form_string))

previous_round_info <- previous_round_info %>%
  select(-predicted_form)

# Get wins in last 5 games for opponent team
for (row in 1:nrow(previous_round_info)) {
  previous_round_info[row, "opponent_form"] <- fetch_ladder(2023, round_number = previous_round_info[row, "round_no"]) %>%
    filter(team.abbreviation == previous_round_info[row, "away_team"]) %>% 
    select(form)
}

for (row in 1:nrow(previous_round_info)) {
  previous_round_info[row, "opponent_form"] = previous_round_info[row, "opponent_form"] %>% 
    substrRight(5)
}

previous_round_info <- previous_round_info %>%
  mutate(opponent_wins_in_last_5 = map_chr(opponent_form, parse_form_string))

previous_round_info <- previous_round_info %>%
  select(-opponent_form)

previous_round_info$wins_in_last_5 <- 
  as.numeric(previous_round_info$wins_in_last_5)
previous_round_info$opponent_wins_in_last_5 <- 
  as.numeric(previous_round_info$opponent_wins_in_last_5)

# Get all player stats for the season up to round 16
season_stats <- data.frame()
for (round in 1:16) {
  round_stats <- fetch_player_stats(season = 2023, round_number = round)
  season_stats <- bind_rows(season_stats, round_stats)
}

# Get sum of ranking points of lineup from each home player's last 5 games
for (row in 1:nrow(previous_round_info)) {
  home_lineup <- fetch_lineup(2023, previous_round_info[row, "round_no"]) %>%
    filter(teamAbbr == previous_round_info[row, "home_team"])
  
  home_rating <- 0
  
  for (prow in 1:nrow(home_lineup)) { # iterate over each player
    if (home_lineup[prow, "position"] %>% pull() != "EMERG") {
      player <- home_lineup[prow, ]
    
    pstats <- season_stats %>% 
      filter(player.player.player.playerId ==
               player[1, "player.playerId"] %>% pull()) %>% 
      tail(3)
    
    home_rating <- home_rating + sum(pstats$ratingPoints)
    }
  }
  
  previous_round_info[row, "home_rating"] <- home_rating
}

# Get sum of ranking points of lineup from each away player's last 5 games
for (row in 1:nrow(previous_round_info)) {
  away_lineup <- fetch_lineup(2023, previous_round_info[row, "round_no"]) %>%
    filter(teamAbbr == previous_round_info[row, "away_team"])
  
  away_rating <- 0
  
  for (prow in 1:nrow(away_lineup)) { # iterate over each player
    if (away_lineup[prow, "position"] %>% pull() != "EMERG") {
      player <- away_lineup[prow, ]
  
      pstats <- season_stats %>% 
        filter(player.player.player.playerId ==
                 player[1, "player.playerId"] %>% pull()) %>% 
        tail(3)
      
      away_rating <- away_rating + sum(pstats$ratingPoints)
    }
  }
  
  previous_round_info[row, "away_rating"] <- away_rating
}
```

Estimate parameters of multiple linear regression model:
```{r}
model <- lm(margin ~ home_position + away_position + last_match_margin + 
              wins_in_last_5 + opponent_wins_in_last_5 + home_rating + 
              away_rating, data = previous_round_info)
summary(model)
```

Assemble predictor data for this match from perspective of home team:
```{r}
# Fetch ladder positions
ladder_positions <- fetch_ladder() %>%
  select(team.abbreviation, position)

# Bind ladder positions with chosen_match_info
chosen_match_info <- chosen_match_info %>%
  left_join(ladder_positions, by = c("home_team" = "team.abbreviation")) %>%
  rename(home_position = position) %>%
  left_join(ladder_positions, by = c("away_team" = "team.abbreviation")) %>%
  rename(away_position = position)

# Fetch scores of last match-up between teams
season_matchup_info <- fetch_results() %>%  arrange(desc(row_number())) %>%
  filter(match.homeTeam.abbr == chosen_match_info[1, "home_team"] %>% pull |
         match.homeTeam.abbr == chosen_match_info[1, "away_team"] %>% pull,
         match.awayTeam.abbr == chosen_match_info[1, "away_team"] %>% pull |
         match.awayTeam.abbr == chosen_match_info[1, "home_team"] %>% pull) %>% 
  select(homeTeamScore.matchScore.totalScore,
         awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)

last_season_matchup_info <- fetch_results(2022) %>% arrange(desc(row_number())) %>%
  filter(match.homeTeam.abbr == chosen_match_info[1, "home_team"] %>% pull |
         match.homeTeam.abbr == chosen_match_info[1, "away_team"] %>% pull,
         match.awayTeam.abbr == chosen_match_info[1, "away_team"] %>% pull |
         match.awayTeam.abbr == chosen_match_info[1, "home_team"] %>% pull) %>% 
  select(homeTeamScore.matchScore.totalScore,
         awayTeamScore.matchScore.totalScore, match.homeTeam.abbr)
  
matchup_info <- bind_rows(season_matchup_info, last_season_matchup_info)
  
if (chosen_match_info[1, "home_team"] == matchup_info[1, "match.homeTeam.abbr"]) {
  chosen_match_info[1, "previous_match_home_score"] <- matchup_info[1, "homeTeamScore.matchScore.totalScore"]
  chosen_match_info[1, "previous_match_away_score"] <- matchup_info[1, "awayTeamScore.matchScore.totalScore"]
} else {
  chosen_match_info[1, "previous_match_home_score"] <- matchup_info[1, "awayTeamScore.matchScore.totalScore"]
  chosen_match_info[1, "previous_match_away_score"] <- matchup_info[1, "homeTeamScore.matchScore.totalScore"]
}

# Get margin from total scores
chosen_match_info <- chosen_match_info %>%
  mutate(last_match_margin = previous_match_home_score - previous_match_away_score)

# Get wins in last 5 games for home team
chosen_match_info[1, "home_form"] <- fetch_ladder(2023) %>%
  filter(team.abbreviation == chosen_match_info[1, "home_team"] %>% pull()) %>% 
  select(form)

chosen_match_info[1, "home_form"] = chosen_match_info[1, "home_form"] %>% 
  substrRight(5)

chosen_match_info <- chosen_match_info %>%
  mutate(wins_in_last_5 = map_chr(home_form, parse_form_string))

# Get wins in last 5 games for away team
chosen_match_info[1, "away_form"] <- fetch_ladder(2023) %>%
  filter(team.abbreviation == chosen_match_info[1, "away_team"] %>% pull()) %>% 
  select(form)

chosen_match_info[1, "away_form"] = chosen_match_info[1, "away_form"] %>% 
  substrRight(5)

chosen_match_info <- chosen_match_info %>%
  mutate(opponent_wins_in_last_5 = map_chr(away_form, parse_form_string))

chosen_match_info <- chosen_match_info %>%
  select(-previous_match_home_score, -previous_match_away_score, 
         -home_form, -away_form)

chosen_match_info$wins_in_last_5 <- 
  as.numeric(chosen_match_info$wins_in_last_5)
chosen_match_info$opponent_wins_in_last_5 <- 
  as.numeric(chosen_match_info$opponent_wins_in_last_5)

# Get all player stats for the season up to round 16
season_stats <- data.frame()
for (round in 1:16) {
  round_stats <- fetch_player_stats(season = 2023, round_number = round)
  season_stats <- bind_rows(season_stats, round_stats)
}

# Get sum of ranking points of lineup from each home player's last 5 games
home_lineup <- fetch_lineup(2023, 16) %>%
  filter(teamAbbr == chosen_match_info[1, "home_team"] %>% pull())

home_rating <- 0

for (prow in 1:nrow(home_lineup)) { # iterate over each player
  if (home_lineup[prow, "position"] %>% pull() != "EMERG") {
    player <- home_lineup[prow, ]
    
    pstats <- season_stats %>% 
      filter(player.player.player.playerId ==
               player[1, "player.playerId"] %>% pull()) %>% 
      tail(3)
    
    home_rating <- home_rating + sum(pstats$ratingPoints)
  }
}
  
chosen_match_info[1, "home_rating"] <- home_rating

# Get sum of ranking points of lineup from each away player's last 5 games

away_lineup <- fetch_lineup(2023, 16) %>%
  filter(teamAbbr == chosen_match_info[1, "away_team"] %>% pull())

away_rating <- 0

for (prow in 1:nrow(away_lineup)) { # iterate over each player
  if (away_lineup[prow, "position"] %>% pull() != "EMERG") {
    player <- away_lineup[prow, ]
  
    pstats <- season_stats %>% 
      filter(player.player.player.playerId ==
              player[1, "player.playerId"] %>% pull()) %>% 
      tail(3)
      
    away_rating <- away_rating + sum(pstats$ratingPoints)
  }
}
  
chosen_match_info[1, "away_rating"] <- away_rating

```

Predict margin using model:
```{r}
predict(model, newdata = chosen_match_info)
```

This margin should be interpreted as an expected Geelong win by 48.07 points.

Here are some residual plots of the model against the training data:
```{r}
# Calculate residuals
res <- resid(model)

# Create a new plot with customized settings
par(mfrow = c(1, 2))  # Divide the plot into two columns

# Plot residuals vs. fitted values
plot(fitted(model), res,
     main = "Residuals vs. Fitted",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 20,  # Set point character to filled circles
     col = "blue"  # Set point color to blue
)

# Add a horizontal line at y = 0
abline(h = 0, col = "red", lty = 2)  # Use a dashed red line

# Plot a histogram of residuals
hist(res,
     main = "Histogram of Residuals",
     xlab = "Residuals",
     ylab = "Frequency",
     col = "lightblue",  # Set histogram color to light blue
     border = "white",   # Set border color to white
     breaks = "FD"       # Use the Freedman-Diaconis rule for bin width
)

# Add a normal curve to the histogram
curve(dnorm(x, mean = mean(res), sd = sd(res)),
      add = TRUE,
      col = "darkblue",
      lwd = 2
)
```

